# Predicting personal activity
## Abstract
The goal of this project is to predict the manner in which they did the exercise.
## Cleaning data
first we load the data. the "#DIV/0!","NA" and empty texts are considered as missing values.

```{r, echo=F, results="hide", cache=T}
library(caret)
```
```{r, cache=T}
## Load the data
training <- read.csv("pml-training.csv", na.strings=c("","#DIV/0!","NA"),
                     stringsAsFactors = F)
testing <- read.csv("pml-testing.csv", na.strings=c("","#DIV/0!","NA"),
                    stringsAsFactors = F)
training$classe <- factor(training$classe)
```
```{r}
## look at the data
dim(training)
table(training$classe)
dim(testing)
```
variables that we know they are irrelevent for prediction will be removed. for example username does not appear to give much information about exercise habits of a person. every single change in features of training set, must be applied to testing set too.
```{r, results="hide"}
## Remove useless features
training <- training[, -which(names(training) == "X")] ## row numbers
training <- training[, -which(names(training) == "user_name")]
testing <- testing[, -which(names(testing) == "user_name" |
                            names(testing) == "X")]

## Remove features with character class
training <- training[, -which(names(training) == "new_window")]
training <- training[, -which(names(training) == "cvtd_timestamp")]
testing <- testing[, -which(names(testing) == "new_window" |
                            names(testing) == "cvtd_timestamp")]
```
all variables that has at least one NA element will be removed.
```{r, results="hide"}
## Remove features with NA
hasNa <- sapply(training, function(x) sum(is.na(x))>0)
training <- training[, !hasNa]
testing <- testing[, !hasNa]
```
to continue
```{r}
##
inTraining <- createDataPartition(training$classe, p = 0.75, list = FALSE)
train <- training[inTraining, ]
test <- training[-inTraining, ]
```
principal component analysis
```{r}
## choose 15 principal components
#preproc <- preProcess(train[,-ncol(train)], method='pca', pcaComp=15)
#trainPc <- predict(preproc, train[,-ncol(train)])
#trainPc <- cbind(trainPc, classe=train$classe)
#preproc <- preProcess(test[,-ncol(test)], method='pca', pcaComp=15)
#testPc <- predict(preproc, test[,-ncol(test)])
#testPc <- cbind(testPc, classe=test$classe)
```
create the model with random forest algorithm.
```{r, cache=T, results="hide"}
## create the model
fitControl <- trainControl(method = "cv", number = 3)
fit <- train(classe ~ ., method="rf", data=train)
```
here we can know how well we predict the `classe` with our model
```{r}
fit$finalModel
predictions <- predict(fit, newdata=test)
confusionMatrix(predictions, test$classe)
table(predictions == test$classe)
```
we predict the final answers which is the real predictions for real test data set
```{r}
answers <- predict(fit, testing)
```
